### 2016-03-02 Draft.tex
Comments to self inserted.

The framing of our model in terms of time-to-detection leads to two practical differences vis-a-vis constant-detection models.  First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}.  Most existing studies have instead constructed regression models on the logit of the equal-interval detection probability \adam{Refs}.  Such an approach is not possible when detection rates are not constant.  Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.  Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection models can only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.

@@

Application of TTDD modeling within the N-mixture framework should be straight-forward for other time-to-detection datasets, such as full detection-history and double-observer.  The key difference will be in the data model, in that responses will take a different form than multinomially distributed first times to detection.  But the underlying abundance and detection models remain the same.

The time-to-detection approach in our model is a departure from the probability-of-detection approach that has been used in distance-removal modeling \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}.  In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.  I'm  beginning to think this may have too-numerous antecedents to cite... in Reidy, the earliest reference is Marsh and Sinclair 1989.  See also McCallum 2005.}.  In avian point-counts, because the overwhelming majority of detections are auditory, $P_a$ is understood to be the probability that a bird vocalizes during the observation period.  Meanwhile, perceptibility is treated as a function of distance $d$: $P_b = g(d)$, with nearby available individuals being detected with higher probability than distant individuals.  The probability of detection is thus formulated as availability times perceptibility: $p^{(det)} = P_aP_b$.  Multiple studies have demonstrated that failure to incorporate distance leads to systemic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.

>>>>

The framing of our model in terms of time-to-detection leads to two practical differences vis-a-vis constant-detection models.  First, in order to model covariate and random effects on detection, we perform mixed effects linear regression on the log of the rate parameter as in \citet{Solymos2013}.  Most existing studies have instead constructed regression models on the logit of the equal-interval detection probability \adam{Amundson2014}.  Such an approach is not possible when detection rates are not constant.  \adam{This statement is false.  It has been done, but the form of the non-constancy must be proscribed.  See Schnute 1983, Scruton and Gibson (ref: Mantyniemi), and Mantyniemi2015 for examples}.  Second, because we can obtain interval-specific detection probabilities from the TTDD by partitioning its cdf, we can directly model the data according to their existing interval structure rather than subdividing the observation period into intervals of equal duration.  Indeed, with a few simplifications, our model fits exact time-to-detection data, whereas existing constant-detection removal models only approximate exact data by subdividing the observation interval into a large number of fine intervals \citep{Reidy2011, Amundson2014}.

@@

Application of TTDD modeling within the N-mixture framework should be straight-forward for other time-to-detection datasets, such as full detection-history and double-observer.  The key difference will be in the data model, in that responses will take a different form than multinomially distributed first times to detection.  But the underlying abundance and detection models remain the same.  \adam{Time-heterogeneous models have been done for mark-recapture (Farcomeni and Scacciatelli 2013, who reference Hwang and Chao 2002).}

The time-to-detection approach in our model is a departure from the probability-of-detection approach that has been used in distance-removal modeling \citep{Farnsworth2005, Diefenbach2007, Solymos2013, Amundson2014}\adam{Alldredge2007?}.  In a distance-sampling context, the probability of detection for any individual is framed as the joint probability of two independent detection events: (i) availability for detection, which occurs during an observation period with probability $P_a$, and (ii) detection of the available individual by an observer, $P_b$, also known as perceptibility \citep{Williams2002, Kery2008, Nichols2009}\adam{I haven't read Williams2002 -- cited in Royle2004.  I'm  beginning to think this may have too-numerous antecedents to cite... in Reidy, the earliest reference is Marsh and Sinclair 1989.  See also McCallum 2005.}.  In avian point-counts, because the overwhelming majority of detections are auditory, $P_a$ is understood to be the probability that a bird vocalizes during the observation period.  Meanwhile, perceptibility is treated as a function of distance $d$: $P_b = g(d)$, with nearby available individuals being detected with higher probability than distant individuals.  The probability of detection is thus formulated as availability times perceptibility: $p^{(det)} = P_aP_b$.  Multiple studies have demonstrated that failure to incorporate distance leads to systemic bias in estimates of abundance \citep{EffordDawson2009, Solymos2013}.
